

## ggplot part 1 - Solutions

Make sure ggplot2 library is installed and then loaded correctly!

Hopefully you have saved the data and RMD into the same folder, and created a project file for that folder. If so then the BeanSurvey dataset can be read in without modifying the code below.

```{r}

library(ggplot2)
BeanSurvey<-read.csv("bean_survey.csv")

```



**Exercise 1. Replace each of the instances of "ZZZ" from the code below to produce a bar chart showing the frequencies of the household head gender variable `GENDERHH`**


ggplot(data=ZZZ, aes(x=ZZZ)) + 
  geom_ZZZ()


** Answer

```{r}

ggplot(data=BeanSurvey, aes(x=GENDERHH)) + 
  geom_bar()

```

To solve the question we need to replace three ZZZ - first with the name of the data `BeanSurvey` and then with the name of the variable within the mapping `GENDERHH`. Be very careful to make sure you spell these exactly correct with upper and lower case.

We also need to set the correct geom_ - in this case to show a bar chart of frequencies we use `geom_bar`



**Exercise 2. Identify and fix the error(s) in this code to produce a boxplot of age of household head by village**


ggplot2(data = BEANSURVEY,  aes(x = AgeHH, y = Village) 
  geom_box()


** Answer

```{r}
ggplot(data = BeanSurvey,  aes(x = AGEHH, y = VILLAGE)) +
  geom_boxplot()
```

There are lots of errors!:
* There is no `+` at the end of the first line
* In the second line there are two brackets opened but only one is closed
* The function is `ggplot` not `ggplot2`
* The data has been entered with the incorrect case - `BeanSurvey` not `BeanSurvey`
* The variables have been entered with incorrect cases. `AGEHH` not `AgeHH`; `VILLAGE` not `Village`
* The correct function is `geom_boxplot()`


**Exercise 3. Produce a histogram of land area. Set the bins so that each bin covers a range of 1 acre**

** Answer

We first need to convert the question into the different ggplot elements:
data-> BeanSurvey
aesthetics->x->LANDAREA
geom->histogram

```{r}

ggplot(data=BeanSurvey,aes(x=LANDAREA))+
  geom_histogram()

```
Once we have this we can think about modifying to only cover 1 acre per bin. We need to find the option `binwidth` from within the geom_histogram function

```{r}
ggplot(data=BeanSurvey,aes(x=LANDAREA))+
  geom_histogram(binwidth = 1)
```



**Exercise 4. Make a scatter plot showing the number of adults on the x axis against the number of children on the y axis. Can you see any limitations to this plot?**

** Answer 

data-> BeanSurvey
aesthetics->x->CHILDREN
aesthetics->y->ADULTS
geom->point

```{r}
ggplot(data=BeanSurvey,aes(x=ADULTS,y=CHILDREN))+
  geom_point()
```
A limitation here is that we have lots of overlapping observations. e.g. lots of people in our data have 2 adults and 3 children. But we only see one point. This could be a case where we might want to do a jitter plot instead.

```{r}
ggplot(data=BeanSurvey,aes(x=ADULTS,y=CHILDREN))+
  geom_jitter(width=0.1,height=0.1)
```
The width and height control how much away from the centre the points are allowed to jitter
